---
title: 'Lab 3A: Foundataions for inference - Sampling Disributions'
author: "Charles Brown"
date: "Sunday, March 29, 2015"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    toc: yes
  html_document:
    fig_caption: yes
    keep_md: yes
    toc: yes
---

```{r DisableSciNotation}
options(scipen=999)
```


### Load Data
```{r LoadData}
load(file = url("http://www.openintro.org/stat/data/ames.RData"))
```

***

### Exploratory Stats
```{r ExploreData}
names(ames)
#
area <- ames$Gr.Liv.Area
(xsum <-summary(area))
hist(x = area, xlab = "Area in Sq. Feet", breaks = 100)

```

#### Question 1: Which of the following is false?
A) The distribution of areas of houses in Ames is unimodal and right-skewed.
B) **50% of houses in Ames are smaller than 1,500 square feet.** 
C) The middle 50% of the houses range between approximately 1,130 square feet and 1,740 square feet. 
D) The IQR is approximately 610 square feet. 
E) The smallest house is 334 square feet and the largest is 5,642 square feet. 

* Half of the houses in Ames are less than the median of `r xsum["Median"]` square feet.

__Question 1 Answer: B)__

***

### The unknown sampling distribution.

```{r SamplingDist}
samp0 <-sample(area,50)
samp1 <-sample(area,50)
#
mean(samp1)
summary(samp1)
#
samp_50 <-sample(area,50)
samp_100 <-sample(area,100)
samp_1000 <-sample(area, 1000)
```

#### Question 2: Suppose we took two more samples, one of size 100 and one of size 1000. Which would you think would provide a more accurate estimate of the population mean?  

A) Sample size of 50 
B) Sample size of 100 
C) **Sample size of 1000**

* Mean of Entire Population of `r nrow(area)`: `r mean(area)`.
* Mean of Sample size 50: `r mean(samp_50)`. 
* Mean of Sample size 100: `r mean(samp_100)`.
* Mean of sample size 1000: `r mean(samp_1000)`.

__Question 2 Answer: The larger the sample size, the better the estimate of the mean.__


The distribution of sample means, called the sampling distribution, can help us understand this variability. In this lab, because we have access to the population, we can build up the sampling distribution for the sample mean by repeating the above steps many times. Here we will generate 5000 samples and compute the sample mean of each.

```{r SamplingDistPlot}
sample_means50 <- rep(NA, 5000)

for(i in 1:5000){
   samp <- sample(area, 50)
   sample_means50[i] <- mean(samp)
   }

hist(sample_means50)

hist(sample_means50, breaks = 50)

summary(sample_means50)
```

#### Exercise: Describe the sampling distribution (the distribution of the sample means that you just created), and be sure to specifically note its center.  


__Excercise Answer: The sampling distribution looks a like a normal distribution with a mean of `r mean(sample_means50)`.__


#### Exercise: To make sure you understand what you’ve done in this loop, try running a smaller version. Initialize a vector of 100 NAs called sample_means_small. Run a loop that takes a sample of size 50 from area and stores the sample mean in sample_means_small. Print the output to your screen (type sample_means_small into the console and press enter).

```{r forlooptest}

sample_means_small <-rep(NA, 100)

for (i in 1: 100)
{
     samp <-sample(area,50)
     sample_means_small[i] <-mean(samp)
    
}

hist(sample_means_small, breaks = 25)
sample_means_small
```

#### Question 3: How many elements are there in this object called sample_means_small? 

A) 0 
B) 30 
C) 50 
D) **100** 
E) 5,000 

__Question 3 Answer: `sample_means_small contains` `r length(sample_means_small)` elements.__


#### Question 4: Which of the following is true about the elements in the sampling distributions you created? 

A) **Each element represents a mean square footage from a simple random sample of 50 houses.** 
B) Each element represents the square footage of a house. 
C) Each element represents the true population mean of square footage of houses. 

__Question 4 Answer: Each element of this sampling distribution vector is a mean of a random sample (size 50) from the population.__


### Sample size and the sampling distribution

To get a sense of the effect that sample size has on our sampling distribution, let’s build up two more sampling distributions: one based on a sample size of 10 and another based on a sample size of 100.

```{r SampleSizeTest}
sample_means10 <- rep(NA, 5000)
sample_means100 <- rep(NA, 5000)

for(i in 1:5000){
  samp <- sample(area, 10)
  sample_means10[i] <- mean(samp)
  samp <- sample(area, 100)
  sample_means100[i] <- mean(samp)
}
```


To see the effect that different sample sizes have on the sampling distribution, plot the three distributions on top of one another.

```{r PloSampleSizeText}
par(mfrow = c(3, 1))

xlimits = range(sample_means10)

hist(sample_means10, breaks = 20, xlim = xlimits)
hist(sample_means50, breaks = 20, xlim = xlimits)
hist(sample_means100, breaks = 20, xlim = xlimits)
```

#### Question 5: It makes intuitive sense that as the sample size increases, the center of the sampling distribution becomes a more reliable estimate for the true population mean. Also as the sample size increases, the variability of the sampling distribution ________. 

A) **decreases**
B) increases 
C) stays the same 

__Answer Question 5: The panel plot shows that the variability decreases as the sample size increases.__


### Now you’ll try to estimate the mean home price.

#### Exercise: Take a random sample of size 50 from price. Using this sample, what is your best point estimate of the population mean?

```{r RandomSamp50Price}
price <- ames$SalePrice
samp <-sample(price,50)
(xsum <-summary(samp))
hist(samp/(1000), xlab = "Price in $Thousands", breaks = 50)
```

__Exercise Answer: The point estimate of the mean is `r xsum["Mean"]/1000`K.__

#### Exercise: Since you have access to the population, simulate the sampling distribution for x¯price by taking 5000 samples from the population of size 50 and computing 5000 sample means. Store these means in a vector called sample_means50. Plot the data, then describe the shape of this sampling distribution. Based on this sampling distribution, what would you guess the mean home price of the population to be?

```{r CreateSampleMeans50DistPrice}

sample_means50 <-rep(NA,5000)

for (i in 1:5000)
{
     samp <-sample(price, 50)
     sample_means50[i] <-mean(samp)
}

hist(sample_means50, breaks = 50)
(xsum <-summary(sample_means50))

```

__Exercise Answer: The shape of the price sampling distribution is normal because the median home price of `r xsum["Median"]/1000`K is approximately the same as the mean home price of `r xsum["Mean"]/1000`K.__


#### Exercise: Change your sample size from 50 to 150, then compute the sampling distribution using the same method as above, and store these means in a new vector called sample_means150. Describe the shape of this sampling distribution, and compare it to the sampling distribution for a sample size of 50. Based on this sampling distribution, what would you guess to be the mean sale price of homes in Ames?

```{r CreateSampleMeans150DistPrice}
sample_means150 <-rep(NA,5000)

for (i in 1:5000)
{
     samp <-sample(price, 150)
     sample_means150[i] <-mean(samp)
}

hist(sample_means150, breaks = 50)
(xsum150 <-summary(sample_means150))

par(mfrow = c(2,1))

xlimits50 <-range(sample_means50)

hist(sample_means50, breaks = 50, xlim = xlimits50)
hist(sample_means150, breaks = 50, xlim = xlimits50)

```

* Sample Size 150 Summary:
```{r summarySampleSize150Price, echo=FALSE}
xsum150
```

* Sample Size 50 Summary:
```{r SummarySampleSizeOf50Price, echo=FALSE}
xsum
```

__Exercise Answer: Both sampling distributions have a near normal shape. The estimated mean price of homes in Ames is $`r mean(c(xsum["Mean"],xsum150["Mean"]))`.__


#### Question 6: Which of the following is false? 

A) **The variability of the sampling distribution with the smaller sample size (sample_means50) is smaller than the variability of the sampling distribution with the larger sample size (sample_means150).** 
B) The means for the two sampling distribtuions are roughly similar.
C) Both sampling distributions are symmetric. 


__Question 6 Answer: A is false. Sample sizes of 150 typically have a smaller spread (`r xsum150["Max."] -xsum150["Min."]`) than sample sizes of 50 (`r xsum["Max."] -xsum["Min."]`).__

***

